<!---
Version: "2013-10-23"
Copyright (C) 2013 Thomas Nauss, GNU GPL (>=3)
-->

The following plotting examples are the lattice version of the above generic 
plotting functions and are also pimped where necessary.

The underlaying example data is again taken from our combined natural disaster and 
Ebola data set which is loaded in the background into the data frame called df. 
We will generally not interpret the plots but just look at formal aspects.

```{r, echo=FALSE, warning=FALSE}
df <- read.table(
  #"D:/active/moc/dm/examples/data_procd/em-dat_disasters_1900-2014_sorted.csv", 
  "/home/ede/tappelhans/uni/marburg/lehre/2014/ws/msc_dm/data/em-dat_disasters_1900-2014_sorted.csv", 
  header = TRUE, sep = ",")
# str(df)
```


##### Multiple plots in one
In the following example, we will again visualize the number of totally affected 
people against the time line. The aggregation looks like this:
```{r}
library(latticeExtra)
dfa <- aggregate(df$total_affected, by = list(df$disaster.group, df$year), FUN = sum)
colnames(dfa) <- c("disaster.group", "year", "total_affected")
```

The lattice version of our previous example looks like this:
```{r}
xyplot(dfa$total_affected ~ dfa$year, 
     xlab = "Year", ylab = "Affected persons",
     main = "Log of persons affected by natural disasters from 1900 to 2013",
     col = "red", fill = "blue", pch = 21,
     scales = list(y = list(log = 10)),
     yscale.components = yscale.components.log10ticks)
```

Let's plot the same data but this time using different colors for the different
desaster groups. To do so, we first define a variable containing the individual
disaster group values:
```{r}
groups <- unique(dfa$disaster.group)
groups
```
As one can see, we have three different groups. Hence, we need three different 
colors which we define now:
```{r}
colors <- c("red", "green", "blue")
```

In order to recreate the generic base graphics version of the multi-colour plot
with different colours for each desaster type in lattice, we need to approach 
this a little differently. Instead of producing an empty plot and then adding
points, in lattice we create one plot for each sub-group which we then plot
on top of each other in a layered fashion. For this, we first need to create a 
list will all the plotting object needed for the final layered plot. We do this
usng an `lapply()`-loop. Note that we already add a legend (key) to each 
individual plot.

```{r}
plot_list <- lapply(seq(groups), function(i) {
  xyplot(dfa$total_affected[dfa$disaster.group == groups[i]] ~
         dfa$year[dfa$disaster.group == groups[i]],      
       xlab = "Year", ylab = "Affected persons",
       main = "Log of persons affected by natural disasters from 1900 to 2013",
       col = colors[i], fill = colors[i], pch=21,
       scales = list(y = list(log = 10)),
       yscale.components = yscale.components.log10ticks,
       key = list(points = list(pch = rep(21, 3), 
                                col = colors,
                                fill = colors),
                  text = list(labels = as.character(groups)), 
                  x = 0.05, y = 0.95, corner = c(0, 1)))
  })
```

Now we have created a plot for each desaster group, e.g. for the 3rd desaster 
group (Complex Desasters) we could now simply call the 3rd element of our 
plot list:

```{r}
plot_list[[3]]
```

In order to combine these individual plots into one plot, the `latticeExtra` 
package provides the `layer()` function which lets you add layers to existing plots.
However, in order to not have too many `layer()` calls we will use another 
function called `latticeCombineLayer()` which is part of an R package developed
by us at Environmental Informatics Marburg. To install this 
package copy the following code into your console:

```{r, eval = FALSE}
install.packages('devtools')
library(devtools)
install_github("environmentalinformatics-marburg/Rsenal")
```

Now we can load this package and use the function to combine multiple plots 
into one:

```{r}
library(Rsenal)
latticeCombineLayer(plot_list)
```


##### Multiple plots on one page
So far we had one plot per page. In lattice we can use the `|` formula argument
to provide a grouping variable so that the solution looks like this (again using
the customised theme and adjusting the ylabels to reflect the logarithmic nature):

```{r, fig.width = 10, fig.height = 10}
bw_theme <- trellis.par.get()
bw_theme$box.dot$pch <- "|"
bw_theme$box.dot$cex <- 3
bw_theme$box.rectangle$col <- "black"
bw_theme$box.rectangle$lwd <- 2
bw_theme$box.rectangle$fill <- "white"
bw_theme$box.umbrella$lty <- 1
bw_theme$box.umbrella$col <- "black"
bw_theme$plot.symbol$col <- "grey40"
bw_theme$plot.symbol$pch <- "*"
bw_theme$plot.symbol$cex <- 2

ndigits <- nchar(as.character(max(df$total_affected)))
ylabls <- 10^(0:ndigits)

# transform the power of then values above analogous to the data values
ytics <- ylabls^(0.5)^3

df$total_affected_sqrt3 <- df$total_affected^0.5^3
bwp <- bwplot(total_affected_sqrt3 ~ region | disaster.group, data = df,    
              scales = list(x = list(rot = 90), 
                            y = list(at = ytics, labels = ylabls)),
              main = "People affected by different disaster types",
              par.settings = bw_theme, layout = c(1, 3))
print(bwp)
```

Note that by default lattice provides sensible axis labelling out of the box.
So there is no need to adjust these. Also not that y axes are labelled alternatingly
on the left an right side of the plot. This is a matter of choice and personal
preference. Some people do not like this, but I find it rather helpful. In any way,
this can be changed.

Note also, that in the last code chunk we stored our plot in an object. Yes, just
like any vector or any other data type, lattice lates you store plots in objects.
This is the main advantage over the base graphics system, as it lets us quickly 
modify existing plots without the need to re-code everything from scratch. For 
this we can use the `update()` function. So if we wanted to change the colours of 
the boxes, we would simply type:

```{r, fig.width = 10, fig.height = 10}
clrs <- colors(1)[1:length(unique(df$region))]

update(bwp, fill = clrs)
```

