---
output: html_document
runtime: shiny
---

The following examples introduce the `shiny` libary, which can be used for 
interactive web-type applications either as stand alone server app, or as a kind
of Rmarkdown web site or slide presentation. The main difference between Rmarkdown
and shiny is that shiny requires either an active R/RStudio session if it is
running local or a shiny server if it is running online.
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
if(Sys.info()["user"] == "shiny"){
  in_path1 <- "/srv/shiny-server/moc/msc-data-management/data_procd/"
  in_path2 <- "/srv/shiny-server/moc/msc-data-management/data_raw/"
  in_path3 <- "/srv/shiny-server/moc/msc-data-management/data_procd/spatial_data/"
  } else {
    setwd("D:/active/moc/dm/examples/")
    in_path1 <- paste0(getwd(), "/data_procd/")
    in_path2 <- paste0(getwd(), "/data_raw/")
    in_path3 <- paste0(getwd(), "/data_procd/spatial_data/")
    }

packages <- c("rCharts", "reshape2", "raster", "rgdal", 
              "latticeExtra", "caret", "gridExtra")
lapply(packages, library, character.only = TRUE)

wb <- read.table(paste0(in_path1, "wb-db_gnipc_co2_1960-2013.csv"), 
                 header = TRUE, sep = ",")
cntry <- read.table(paste0(in_path2, "wb-db_country_classes.csv"), 
                    header = TRUE, sep = ",")
wbc <- merge(wb, cntry[,!(grepl("Economy|X", colnames(cntry)))], 
             by.x = "Country.Code", by.y = "Code")
wbc$Region[wbc$Region == ".."] <- NA
wbc$Region <- droplevels(wbc$Region)
wbc$co2_log <- log(wbc$co2)
wbc$gni_log <- log(wbc$gni)
wbc <- wbc[complete.cases(wbc),]
load(paste0(in_path3, "dm-es-12_02_rf.RData"))
ndvi_original <- raster$ndvi_fogo_landsat
prediction_original <- predict(raster, model.train[[1]])
clrs.ndvi <- colorRampPalette(brewer.pal(8,"YlGn"))(256)
clrs.change <- colorRampPalette(brewer.pal(11,"Spectral"))(256)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

##### Histogram

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r, echo=FALSE, }
slider <- input_slider(10, 100)
mtcars %>% ggvis(~wt, ~mpg) %>%
  layer_points(size := slider) %>% 
  layer_points(fill := "red", opacity := 0.5, size := slider)

mtcars %>% ggvis(~wt, ~mpg) %>%
  layer_points(size := input_slider(100, 1000, label = "black")) %>%
  layer_points(fill := "red", size := input_slider(100, 1000, label = "red"))


p1 <- ggvis(~wt, ~mpg, data = mtcars)
p1
layer_points(p1, size := input_slider(100, 1000, label = "black"))
layer_points(p1, fill := "red", size := input_slider(100, 1000, label = "red"))

inputPanel(
  
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2,
              animate = animationOptions(interval = 1000, loop = TRUE, 
                                         playButton = NULL, pauseButton = NULL))
  )

renderPlot({
  names(iris) = gsub("\\.", "", names(iris))
  rPlot(SepalLength ~ SepalWidth | Species, data = iris, 
        color = 'Species', type = 'point')
  })
```
