---
output: html_document
---

The following examples show how to produce histogram, box whisker and scatter
plots using the `iPlots` package. The exemplary data frame contents are based on 
the already known World Bank data sets. Since they only work from inside an 
active R environment, no output can be shown here.

Please note that you need the `rJava` package in order to run iPlots. Depending
on your system configuration it might be a good idea to run a 32bit R version
(and the corresponding Java version) even if you have a 64bit processor.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
wb <- read.table(
  "D:/active/moc/dm/examples/data_procd/wb-db_gnipc_co2_1960-2013.csv", 
  header = TRUE, sep = ",")
cntry <- read.table(
  "D:/active/moc/dm/examples/data_raw/wb-db_country_classes.csv", 
  header = TRUE, sep = ",")
wbc <- merge(wb, cntry[,!(grepl("Economy|X", colnames(cntry)))], 
             by.x = "Country.Code", by.y = "Code")
wbc$Region[wbc$Region == ".."] <- NA
wbc$Region <- droplevels(wbc$Region)
wbc$co2_log <- log(wbc$co2)
wbc$gni_log <- log(wbc$gni)
wbc <- wbc[!(is.na(wbc$Region)),]
```



```{r, warning=FALSE, results="hide"}
library(iplots)
ihist(wbc$gni_log)
ihist(wbc$co2_log)
ibox(wbc$gni_log, wbc$Region)
ibox(wbc$co2_log, wbc$Region)
iplot(wbc$gni_log, wbc$co2_log)


```
