<!---
Version: "2013-10-23"
Copyright (C) 2013 Thomas Nauss, GNU GPL (>=3)
-->

The following plotting examples will revisit R's generic plotting functions
and pimp them up a little bit. We will not go into details since if you really
want to visualize your data as heatmap, use lattice or ggplot2. 

The underlaying example data is taken from our combined natural disaster and 
Ebola data set which is loaded in the background into the data frame called df. 
We will generally not interpret the plots but just look at formal aspects.

```{r, echo=FALSE, warning=FALSE}
df <- read.table(
  #"D:/active/moc/dm/examples/data_procd/em-dat_disasters_1900-2014_sorted.csv", 
  "/home/ede/tappelhans/uni/marburg/lehre/2014/ws/msc_dm/data/em-dat_disasters_1900-2014_sorted.csv", 
  header = TRUE, sep = ",")
# str(df)
```

Prior to plotting the heat map, we need to create a matrix which contains 
the desired information. In our case we want to visualize the number of 
disasters per year per region. To get the data set we can use the table
function applied to column year and region of our original data frame.
The function aligns the regions and years column and row lables and counts the
number of events per year.

After transforming the data using a square root function (for a better value and
hence color stretch in the heat map), we produce it using the `heatmap` function:
```{r}
data <- table(df$year, df$region)
head(data)
data <- data^0.5^2

clrs_hcl2 <- function(n) {
  hcl(h = seq(0, 260, length.out = n), 
      c = 60, l = seq(10, 90, length.out = n), 
      fixup = TRUE)
  }

heatmap(data, Rowv = NA, Colv = NA, margins = c(5,5), 
        scale = "column", cexCol=0.75, col = clrs_hcl2(100))
```
