<!---
Version: "2013-10-23"
Copyright (C) 2013 Thomas Nauss, GNU GPL (>=3)
-->
Functions should be used if certain procedures are recycled within or across
scripts. To define a function, the `fuction` class is used.

##### General structure
A very simpel function must include the class name followed by an empty list
of parameters:
```{r}
simpleFunction <- function(){
  print("I am a simple function")
}
simpleFunction()
```
The last line executes the function and will look very familiar to you since you 
already know many functions like `read.table`, `rowSums` or `if`. Even `+` and 
all other arithmetic operations are functions.

Keep in mind that since R is an interpreter language, you will have to define 
the function above the first call or in another script which is imported in
the begining. The latter is realized by the `source` function. Imagin that the
above function would be in a file called "simpleFunction.R" which is stored in
your working directory. Then you can load this function using the following
statement:

```{r, eval=FALSE}
source("simpleFunction.R")
```

##### Using external variables inside a function
If you want to use a variable defined outside the function within it, one (bad)
option would be to adress it directly. Here's an example (you should not 
remember!):
```{r}
simpleFunction <- function(){
  print("I am a simple function which prints the content of a variable:")
  print(test)
}
test <- "Content of variable test"
simpleFunction()
```

As a general rule, never use anything inside a function from the outside except
you explicitly supply it through the function call. So the above example should
be implemented like the following:
```{r}
simpleFunction <- function(content){
  print("I am a simple function which prints the content of a variable:")
  print(content)
}
test <- "Content of variable test"
simpleFunction(test)
```
In this case, the string stored in variable `test` is given to the variable 
`content` which is used within the function's print statement.

##### Return a variable from a function
A function generally returns whatever the last line executed within the function
would return. Hence, in the above example, the function would return the last 
print statement. Let's proof that first:
```{r}
return_variable <- simpleFunction(test)
return_variable
```
Although this might save a couple of characters of your code, it is not good
practice. Therefore, if you want to return something from a function, 
explicitly use the `return` statement. 

Let's say you have a function which converts character vectors to lower case and 
numeric vectors to the square root. To return the results of this function, one 
implementation could look like like that:
```{r}
conversion <- function(element){
  if(is.character(element)){
    return(tolower(element))
    }    else if(is.numeric(element)){
      return(sqrt(element))
      }
  }
```
Let's execute the function:

```{r}
a <- c("A", "B", "C", "A", "B", "A", "A")
c <- c(1, 2, 3, 4, 5, 6, 7)
result <- conversion(a)
result
result <- conversion(c)
result
```
It obviously works. Of course you can also store the result in a variable first
before returning it, e.g. if you want to have an even more clean code structure
or if you want to print or write something in the end.
```{r}
conversion <- function(element){
  if(is.character(element)){
    temp <- tolower(element) 
    print(paste0("Element ", element, " is converted to ", temp, "\n"))
    return(temp)
    }    else if(is.numeric(element)){
      temp <- sqrt(element)
      print(paste0("Element ", element, " is converted to ", temp))
      return(temp)
      }
  }
result <- conversion(a)
result
```
##### Modifying a variable supplied to a function
This is very easy since it is not possible:-). Just for the record:
```{r}
willNotChangeAnOutsideVariableCalledVariable <- function(){
  variable <- variable + 5
  return(variable)
  }
variable <- c(1,2,3,4,5)
willNotChangeAnOutsideVariableCalledVariable()
variable
```
While variables existing outside a function can be used within a function, variables
having the same name as an outside variable only exist inside the function, not
outside.
