---
output: html_document
---


The following examples show some publication quality, map-type visualizations. 
The same vector and raster datasets as in the I/O example are taken. The vector
is already projected to the cartographic reference system of the raster layer.
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
library(raster)
library(sp)
library(rgdal)
library(grid)
library(latticeExtra)
library(RColorBrewer)
library(rworldmap)
library(rworldxtra)

setwd("D:/active/moc/dm/examples/data_procd/spatial_data/")
raster <- raster("LC82100502014328LGN00_B3.tif")
raster_latlon <- raster("LC82100502014328LGN00_B3_latlon.tif")
raster_stack <- stack("LC82100502014328LGN00_B1.tif", 
                      "LC82100502014328LGN00_B2.tif", 
                      "LC82100502014328LGN00_B3.tif")
vector <- readOGR("data_2014_subset1.shp", layer = "data_2014_subset1")
vector_utm <- spTransform(vector, CRS(projection(raster)))
```

#####  Visualization of single raster layers
As alrady introduced in the basic visualization examples from last session, the
`spplot` function which is based on lattice/trellice can be your workhorse in 
producing publication quality maps. Let's start with a simple greyscale
visualization. Please note that this call of spplot actually uses the `levelplot` 
function to draw the raster layer (which will become important later):

```{r}
colorsheme <- gray.colors(256)
spplot(raster, col.regions = colorsheme)
```

The clouds are obvious, but contrast stretch is far from optimal. Using a histogram
stretch gives the following results:

```{r}
breaks <- quantile(raster, seq(0.0, 1.0, length.out = 256))
colorkey_scale <- 1/breaks

colorkey_lable_pos <- pretty(colorkey_scale, n = 4)
colorkey_lables <- round(rev((1/pretty(colorkey_scale, n = 4))[2:5]),0)

spplot(raster, col.regions = colorsheme, at = breaks, 
       colorkey = list(at = as.numeric(colorkey_scale), 
                       labels = list(lables = colorkey_lables, 
                                     at  = colorkey_lable_pos)))
```

Better but rather to light. Before we go on, let's have a quick look on the code.
Variable breaks is used to store the values of 256 quantiles which are equaly 
distributed between 0 and 1. This basically leads to a histogram stretch of the
data values since the value differences between the quantiles are smaller for 
ranges with frequent values. 

Since we want an undistored color scale, the breaks
are inverted (variable colorkey_scale) and used to define positions of ticks and
labels of the color key. The `pretty` function is used for the tick locations 
since it divides a given value range in about the desired number n of sequences.

The lables for the ticks are finally computed by inverting the colorkey_scale values
back to the original data values. Since the largest original value had become the
smallest inverted value, the sequence is reversed by the `rev` function. To
get rid of needlees value fractions, the labels are rounded to full integer values.

Finally, the raster layer is passed to the `spplot` function along with the 
location of the breacks and the information for the colorkey.

An alternative contrast stretch method would be e.g. a standard deviation based
approach. Therefore, one could use the mean value of the data +/- one (or more)
standard deviations. The following code should be self explaining after the
above example. Since it is a linear stretch, the labeling transformation/scaling
is not necessary which makes it a lot easier:

```{r}
min <- max(mean(getValues(raster)) - sd(getValues(raster)), 0)
max <- mean(getValues(raster)) + sd(getValues(raster))

breaks <- seq(min, max, length.out = 256)

spplot(raster, col.regions = gray.colors(256), at = breaks)
```

If you think that a map-type figure should have a cartographic grid, there you go:

```{r}
yat = seq(extent(raster)@ymin, 
                extent(raster)@ymax, length.out = 5)
xat = seq(extent(raster)@xmin, 
                extent(raster)@xmax, length.out = 5)

spplot(raster, col.regions = gray.colors(256), at = breaks,
       panel = function(...){
         panel.levelplot(...)
         panel.abline(h = yat, v = xat, col = "grey0", lwd = 0.8, lty = 3) 
         },
       scales = list(x = list(at = xat),
                     y = list(at = yat)))
```

The grid is just a set of vertical/horizontal lines which positions have been
defined based on the x and y (i.e. easting and northing) extends of the spatial
data layer. In the example above, 5 equally spaced lines in x and y direction
are defined by using the `seq` function with the minimum x/y value as starting 
and the maximum x/y value as end of the value range. The length.out attribute
defines the number of values in the returned sequence.

Since we have to plot two elements now, the `spplot` function is extended by the
panel argument. The `panel.levelplot` function draws the levelplot, i.e. the 
raster layer since `levelplot` is the actual plotting routine of `spplot` if a
raster layer is passed to the function.

The grid on the other hand is just a set of simple lines which can be drawen by
e.g. the `abline` function. The `panel.abline` function is just the panel
equivalent to it. Paramters h and v define the locations of horizontal and 
vertical lines and the rest of the function is self explaining.

The scales finally takes care of the labeling of the grid lines. Withouth it, the
lines would just appear without the northing/easting values.

Just for the sake of completness and to make the preparation for the `spplot` 
function it even more confusing, the following example shows a visualization of
a geographic projection. Therefore, the raster dataset is projected to a geographic
coordinate system (variable raster_latlon):

```{r}
yat_latlon = round(seq(extent(raster_latlon)@ymin, 
                extent(raster_latlon)@ymax, length.out = 5), 2)
ylabs_latlon = ifelse(yat_latlon < 0, paste(yat_latlon, "°S", sep = ""),
               ifelse(yat_latlon > 0, paste(yat_latlon, "°N", sep = ""),
                      paste(yat_latlon, "°", sep = "")))

xat_latlon = round(seq(extent(raster_latlon)@xmin, 
                extent(raster_latlon)@xmax, length.out = 5), 2)
xlabs_latlon = ifelse(xat_latlon < 0, paste(xat_latlon, "°W", sep = ""),
               ifelse(xat_latlon > 0, paste(xat_latlon, "°E", sep = ""),
                      paste(xat_latlon, "°", sep = "")))

spplot(raster_latlon, col.regions = gray.colors(256), at = breaks,
       panel = function(...){
         #          grid.rect(gp=gpar(col=NA, fill="grey50"))
         panel.levelplot(...)
         panel.abline(h = yat_latlon, v = xat_latlon, col = "grey0", lwd = 0.8, lty = 3) 
         },
       scales = list(x = list(at = xat_latlon, labels = xlabs_latlon),
                     y = list(at = yat_latlon, labels = ylabs_latlon)))
```

Basically, the code above shows nothing new for the `spplot` function but the 
grid is now additionally labled with the characters N and W. This is possible since
the lables are explicitly provided to the scales attribute of the function.

As a last example for a single raster layer, we will add country boundaries on 
top of the raster image. Therefore, the `getMaps` function of the rworldmap 
package is used. Since it returns the data in geographical coordinates, we use
the extend of the projected raster layer from above to crop the world wide
country boundaries to the desired region and project the data to the UTM projection
of the raster layer afterwards by using the `spTransform` function.

```{r}
ext <- extent(raster_latlon)
map <- getMap(resolution = "high")
map <- crop(map, ext)
map <- spTransform(map, CRS(projection(raster)))

ext <- extent(raster)
spplot(raster, col.regions = gray.colors(256), at = breaks,
       xlim=c(ext@xmin, ext@xmax), ylim=c(ext@ymin, ext@ymax),
       panel = function(...){
         panel.levelplot(...)
         panel.abline(h = yat, v = xat, col = "grey0", lwd = 0.8, lty = 3) 
         sp.polygons(map, lwd = 0.5)
         },
       scales = list(x = list(at = xat),
                     y = list(at = yat)))
```

As you might notice, the country vector does not match the island. This is an 
error in the data set but just look at it as a how to recipe.

#####  Visualization of raster and vector layers
Visualizing the location of e.g. points of a vector layer is easy and straight
forward. Again, the panel attribute is used but instead of using a panel for 
lines, the vector points are visualized by using the `sp.points` panel function:

```{r}
spplot(raster, col.regions = gray.colors(256), at = breaks,
       panel = function(...){
         panel.levelplot(...)
         sp.points(vector_utm) 
         },
       scales = list(x = list(at = xat),
                     y = list(at = yat)))
```

So much for the easy part. The call of the `spplot` function becomes more 
complicated, if one wants to visualize a specific attribute of the vector along
with an appropriate legend. In the following example, the vegetation coverage
should be visualized:

```{r}
vector_classes <- cut(vector_utm@data$COVRG, c(0, 20, 40, 60, 80, 100, 120))
vector_colors <- colorRampPalette(brewer.pal(6,"Greens"))(6)

plt <- spplot(raster, col.regions = gray.colors(256), at = breaks,
       key = list(space = 'left', text = list(levels(vector_classes)), 
                     points = list(pch = 21, cex = 2, fill = vector_colors)),
       colorkey=list(space="right"),
       panel = function(...){
         panel.levelplot(...)
         panel.abline(h = yat, v = xat, col = "grey0", lwd = 0.8, lty = 3) 
         },
       scales = list(x = list(at = xat),
                     y = list(at = yat)))

orl <- spplot(vector_utm, zcol = "COVRG", col.regions = vector_colors, 
              cuts = c(0, 20, 40, 60, 80, 100, 120))

plt + as.layer(orl)
```

Let's have a look at the code above. The first `spplot` call looks like the ones
above except for two aspects. First, the colorkey argument is used to specify the
location of the default color legend bar. Since it is set to the right hand side
of the plot which is the default, it is actually not necessary, but for 
documentation purposes we include it anyway. 

The second and important difference is the key argument. It defines the legend for
the vegetation coverage attribute of the vector dataset. The legend is set to the
left hand side of the plot (argument space) and the labels are based on the 
pre-defined classes into which the attribute information has been divided by the
`cut` function in the first line of the code example. The points argument 
finally defines the size and type of the points along with the colors.

As a result, a map-type visualization of the raster layer along with both the
legend for the raster values and the vector values will be stored in the variable
plt.

The vector layer attribute is visualized by a seperate call to the `spplot` function. 

The last line combines both plots in order to visualize the final map.

For those who really would like to have a continous color scale bar for the
vector data set, the following code is of interest. But if you feel dizzy 
already from the example above, do not have a look at it now.

```{r}
colorsheme <- colorRampPalette(brewer.pal(9,"Greens"))(256)

vector_classes <- cut(vector_utm@data$COVRG, seq(0, 120, length.out = 256))

plt <- spplot(raster, col.regions = gray.colors(256), at = breaks,
       colorkey=list(space="right"),
       panel = function(...){
         panel.levelplot(...)
         panel.abline(h = yat, v = xat, col = "grey0", lwd = 0.8, lty = 3) 
         },
       scales = list(x = list(at = xat),
                     y = list(at = yat)))

orl <- spplot(vector_utm, zcol = "COVRG", col.regions = colorsheme, cuts = 256,
              colorkey = TRUE)

comb <- plt + as.layer(orl)
print(comb)

downViewport(trellis.vpname(name = "figure")) 
vp1 <- viewport(x = -0.25, y = 1,
                height = 1, width = 1,
                just = c("centre", "top"),
                name = "key.vp") 
pushViewport(vp1) 
draw.colorkey(key = list(col = colorsheme, width = 2,
                          at = seq(0, 110, length.out = 256),
                          space = "left"), draw = TRUE)
```

Basically, the raster and vector layer is plotted independently from each other
and this time, the legend of the vector layer is not included in the call
of the `spplot` function for the raster layer.

Although it might look like that, but the colorkey argument used within the 
`spplot` call for the vector layer does not matter, too. In fact, we could have
set it to TRUE in the previous example, too, but it would not have made any 
difference. We just set it to TRUE because in that case, a continoues
color scheme is drawn for the plot which is faster than a color legend with 256
individual circles and values (which would be the result from colorkey = FALSE
which is the default).

The legend for the vector layer is actually drawn after the combined plot has
been printed (i.e. `print(comb)`). Therefore, the vieport (i.e. the region 
which has just been used to draw the figure) is identified and then a new
vieport is added to the existing one at a specific position. In this case,
the position is just left of the main plotting area which is used for the 
visualizatino of the raster and vector layer and which ends at x coordinate 0.

This viewport is finally set as the active one by using the `pusViewport` 
function and the `draw.colorkey` function is used to draw the legend of the
vector attribute at exactly this viewport's location. If you want to know 
more about that kind of stuff, have a look at the `grid` package and the 
respective documentation.


#####  RGB color plot
As a last example, this is how you can produce a RGB image from a raster stack.
The raster stack contains bands 1 through 3 of a Landsat scene:
```{r}
names(raster_stack)
```

Since contrast stretch is also important for color composits, the bands are 
stretched using some interactively defined bounds for the linear stretch:
```{r}
raster_stack[[1]] <- 
  stretch(raster_stack[[1]], minv = 0, maxv = 255, minq = 0.01, maxq = 0.75)
raster_stack[[2]] <- 
  stretch(raster_stack[[2]], minv = 0, maxv = 255, minq = 0.01, maxq = 0.72)
raster_stack[[3]] <- 
  stretch(raster_stack[[3]], minv = 0, maxv = 255, minq = 0.01, maxq = 0.80)
```

Once this is done, the composit can be plotted using the `plotRGB` function.
Since the raster stack contains the bands in blue, green, red order, the 
appropriate stack layer for the individual colors (i.e. r, g, b) has to be 
provided to the function:
```{r}
plotRGB(raster_stack, r = 3, g = 2, b = 1, stretch = "lin")
```
