<!---
Version: "2013-10-23"
Copyright (C) 2013 Thomas Nauss, GNU GPL (>=3)
-->

The following plotting examples are the lattice version of the above generic 
plotting functions and are also pimped where necessary.

The underlaying example data is again taken from our combined natural disaster and 
Ebola data set which is loaded in the background into the data frame called df. 
We will generally not interpret the plots but just look at formal aspects.

```{r, echo=FALSE, warning=FALSE}
df <- read.table(
  "/home/ede/tappelhans/uni/marburg/lehre/2014/ws/msc_dm/data/em-dat_disasters_1900-2014_sorted.csv", 
  header = TRUE, sep = ",")
# str(df)
```

The lattice version of the generic simple boxplot of our square root transformed 
total number of affected can be produced as follows - the equivalent for `boxplot()`
in lattice is `bwplot()`:
```{r}
library(latticeExtra)
df$total_affected_sqrt3 <- df$total_affected^(0.5)^3
bwplot(total_affected_sqrt3 ~ region, data = df, 
       scales = list(x = list(rot = 90)))
```

Note that in order to modify the axes, lattice provides an argument called `scales`.
In general, lattice stores and retrieves pretty much all its settings through 
lists, so that we need to provide a list to the `scales` argument. And furthermore, 
we also need to provide a list of graphical parameter settings for eeach of the 
x and y axes as lists. With this we can reproduce the cube root adjusted y axis labels 
from the generic version produced earlier:
```{r}
# get maximum value as power of 10 (by counting digits of the max value)
# and create a vector of these powers starting with 10^0
ndigits <- nchar(as.character(max(df$total_affected)))
ylabls <- 10^(0:ndigits)
print(ylabls)

# transform the power of then values above analogous to the data values
ytics <- ylabls^(0.5)^3

# plot the boxplot with modified y axis
bwplot(total_affected_sqrt3 ~ region, data = df, 
       scales = list(x = list(rot = 90), 
                     y = list(at = ytics, labels = ylabls)))
```

I don't know about you, but I surely do not like the defauölt look of lattice
boxplots. This is, however, easy to change. As I said before, lattice stores
all its graphical settings in a list. The default list can be assessed using 
a function called `trellis.par.get()`. With this function we can create our own 
modified version of graphical settings and use those when plotting our boxplot:

```{r}
bw_theme <- trellis.par.get()
bw_theme$box.dot$pch <- "|"
bw_theme$box.dot$cex <- 3
bw_theme$box.rectangle$col <- "black"
bw_theme$box.rectangle$lwd <- 2
bw_theme$box.rectangle$fill <- "white"
bw_theme$box.umbrella$lty <- 1
bw_theme$box.umbrella$col <- "black"
bw_theme$plot.symbol$col <- "grey40"
bw_theme$plot.symbol$pch <- "*"
bw_theme$plot.symbol$cex <- 2

bwplot(total_affected_sqrt3 ~ region, data = df, 
       scales = list(x = list(rot = 90), 
                     y = list(at = ytics, labels = ylabls)),
       par.settings = bw_theme)
```

Do you realise how we produce a `bw_theme` object through `trellis.par.get()`
of which we then modify several settings in order to get a much more pleasant
result. This modified graphics setting list (it is a list, you can test using `str`)
is then passed to the `par.settings` argument in the `bwplot()` call.
