---
output: html_document
---


The following examples show how one can access data values of geospatial datasets 
and how basic visualization can be done. The same vector and raster datasets as
in the I/O example are taken.
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
library(raster)
library(sp)
library(rgdal)
library(latticeExtra)

setwd("D:/active/moc/dm/examples/data_procd/spatial_data/")
vector <- readOGR("data_2014_subset1.shp", layer = "data_2014_subset1")
raster <- raster("LC82100502014328LGN00_B5.tif")
```

#####  Accessing values in spatial data frames
To access the attribute values of a spatial vector data frame, one has to 
access the "data" section of the variable. This is done using the @ symbol:
```{r}
str(vector@data)
```
Aside from that, everything else is just as for the data frames. Here are some
examples:
```{r}
head(vector@data$ID)
vector@data[1:3, ]
```
In order to access values of a raster layer, one can use the index values of
rows and columns directly. Again some examples:
```{r}
raster[5, 6]   # Grid/pixel value of row 5 and column 6
raster[5:6, 6:7]   # Grid/pixel value of r5/c6, r5/c7, r6/c5, r6/c7
```
In order to get all values from the raster in the correct geometry, one has
two options:
```{r}
option1 <- raster[,]
option2 <- getValues(raster)
identical(option1, option2)
```

#####  Basic visualization
Basic visualization of a raster or vector geo-dataset is very simple. For raster
data, you can use the `plot` function or the `spplot` function which is based 
on lattice.
```{r}
plot(raster)
spplot(raster)
```
Simple enough but although this example should just illustrate the basic plooting,
let's change at least the color scheme of the ssplot. We use the standard pallet
from R's basic plotting routines:
```{r}
spplot(raster, col.regions = rev(terrain.colors(256)))
```
Now the colors are better (but not honest) but the image does still not look good.
We come back to contrast stretching in the next session.

For vector data, the `spplot` function can be used, too:
```{r}
spplot(vector)
```
The standard settings will produce one plot for each attribute of the dataset.
In addition, an identical value range is visualized over all attributes. To change
that, the zcol argument of the function can be used to select the respective
attribute:
```{r}
spplot(vector, zcol = "NAT")
```
For a continoues range of values, use the colorkey argument:
```{r}
spplot(vector, zcol = "NAT", colorkey = TRUE)
```

In order to overlay a vector on a raster map, both have to have the same projection.
Since our vector data is still in a geographical coordinate system, it has to be
transformed:
```{r}
vector_utm <- spTransform(vector, CRS(projection(raster)))
```
Afterwards
```{r}
plot(raster) 
plot(vector_utm, add = TRUE)
```

```{r}
spplot(raster, col.regions = rev(terrain.colors(256))) + 
  as.layer(spplot(vector_utm, zcol = "NAT", colorkey = TRUE))
```
