---
output: html_document
runtime: shiny
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
if(Sys.info()["user"] == "shiny"){
  in_path1 <- "/srv/shiny-server/moc/msc-data-management/data_procd/"
  in_path2 <- "/srv/shiny-server/moc/msc-data-management/data_raw/"
  in_path3 <- "/srv/shiny-server/moc/msc-data-management/data_procd/spatial_data/"
  } else {
    setwd("D:/active/moc/dm/examples/")
    in_path1 <- paste0(getwd(), "/data_procd/")
    in_path2 <- paste0(getwd(), "/data_raw/")
    in_path3 <- paste0(getwd(), "/data_procd/spatial_data/")
    }

packages <- c("raster", "rgdal", "latticeExtra", "caret", "gridExtra")
lapply(packages, library, character.only = TRUE)

wb <- read.table(paste0(in_path1, "wb-db_gnipc_co2_1960-2013.csv"), 
                 header = TRUE, sep = ",")
cntry <- read.table(paste0(in_path2, "wb-db_country_classes.csv"), 
                    header = TRUE, sep = ",")
wbc <- merge(wb, cntry[,!(grepl("Economy|X", colnames(cntry)))], 
             by.x = "Country.Code", by.y = "Code")
wbc$Region[wbc$Region == ".."] <- NA
wbc$Region <- droplevels(wbc$Region)
wbc$co2_log <- log(wbc$co2)
wbc$gni_log <- log(wbc$gni)
wbc <- wbc[complete.cases(wbc),]
load(paste0(in_path3, "dm-es-12_02_rf.RData"))
ndvi_original <- raster$ndvi_fogo_landsat
prediction_original <- predict(raster, model.train[[1]])
clrs.ndvi <- colorRampPalette(brewer.pal(8,"YlGn"))(256)
clrs.change <- colorRampPalette(brewer.pal(11,"Spectral"))(256)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

##### Histogram

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r, echo=FALSE, }
inputPanel(
  
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2,
              animate = animationOptions(interval = 1000, loop = TRUE, 
                                         playButton = NULL, pauseButton = NULL))
  )

renderPlot({
  hist1 <- histogram(wbc$gni_log, type = "density", 
            breaks = as.numeric(input$n_breaks), col = "white",
       xlab = "Logarithm of GNI", ylab = "Density", 
       main = "World bank data GNI values")

  dens1 <- densityplot(wbc$gni_log, adjust = input$bw_adjust, na.rm = TRUE, 
              col = "darkgreen", lwd = 3.0, plot.points = FALSE)
  
  plot1 <- hist1 + as.layer(dens1)

  hist2 <- histogram(wbc$co2_log, type = "density", 
            breaks = as.numeric(input$n_breaks), col = "white",
       xlab = "Logarithm of CO2", ylab = "Density", 
       main = "World bank data CO2 values")

  dens2 <- densityplot(wbc$co2_log, adjust = input$bw_adjust, na.rm = TRUE, 
              col = "blue", lwd = 3.0, plot.points = FALSE)
  
  plot2 <- hist2 + as.layer(dens2)
  
  grid.arrange(plot1, plot2, ncol = 2)
  })
```


##### XYPlot
```{r, echo=FALSE, }
inputPanel(
  
  selectInput("model_select", label = "Select model:",
              choices = c("Linear model", "Exponential model",
                          "Self starting logistic model", "Tolerance curve"),
              selected = "Linear model")
  )

renderPlot({

  if(input$model_select == "Linear model"){
    # linear model
    l_model <- lm(wbc$co2_log ~ wbc$gni_log)
    thv <- predict.lm(l_model, wbc)
    } 
  else if(input$model_select == "Exponential model"){
    f <- function(y0, x, a, b) {y0 + a * exp(b * x)}
    exp_model <- nls(co2_log ~ f(y0, gni_log, a, b),
                     data = wbc,
                     start = list(y0 = 3, a = 0.5, b = 0.5)) 
    thv <- predict(exp_model, wbc$gni_log)
    } 
  else if(input$model_select == "Self starting logistic model"){
    Asym <- 1; xmid <- 1; scal <- 1
    getInitial(co2_log ~ SSlogis(gni_log, Asym, xmid, scal), data = wbc)
    ssl_model <- nls(co2_log ~ SSlogis(gni_log, Asym, xmid, scal), data = wbc)
    thv <- predict(ssl_model, wbc$gni_log)
    } 
  else if(input$model_select == "Tolerance curve"){
    tc_model= nls(co2_log ~ 1 + a * (gni_log^2) + b * gni_log + c, wbc,
                  start = list(a=1, b=1, c=1))
    thv <- predict(tc_model, wbc$gni_log)
    }
  
  xyplot(wbc$co2_log[wbc$co2_log >  thv] ~ 
           wbc$gni_log[wbc$co2_log >  thv],
         col = "red", ylim = range(wbc$co2_log, na.rm = TRUE),
         xlab = "Logarithm of GNI", ylab = "Logarithm of CO2", 
         main = "World bank data values") +
    as.layer(xyplot(wbc$co2_log[wbc$co2_log <=  thv] ~ 
                      wbc$gni_log[wbc$co2_log <=  thv], 
                    col = "blue")) + 
    as.layer(xyplot(thv ~ wbc$gni_log, type = "spline", 
                    col = "green", lwd = 5))
  })
```



##### Random forest
```{r, echo=FALSE}
inputPanel(
  selectInput("ndvi", label = "Change percentage NDVI:",
              choices = seq(0, 200, 10), selected = 100)
  )

renderPlot({
  raster$ndvi_fogo_landsat <- ndvi_original * (as.numeric(input$ndvi) / 100)
  prediction_raster <- predict(raster, model.train[[1]])

  
  plot1 <- spplot(prediction_raster,col.regions = clrs.ndvi,
                  colorkey = list(at = seq(0, 11, length.out = 256)),
         main = "Predicted species richness \n based on Landsat 8 data",
         sub = paste0("Result is based on ",input$ndvi, "% of 2014 NDVI values"),
         scales = list(draw = TRUE))
  change_raster = (1-prediction_raster / prediction_original) * 100
  plot2 <- spplot(change_raster,col.regions = clrs.change, 
                  colorkey = list(at = seq(-100, 100, length.out = 256)),
                  main = "Percentage change in prediction \n results compared to 100% NDVI",
                  scales = list(draw = TRUE))
  grid.arrange(plot1, plot2, ncol = 2)
  })
```

