---
output: html_document
runtime: shiny
---

The following examples show some interactive visualizations using the `shiny`
library in the markdown version. The already known CO2 and GNI data from the
World Bank Group as well as the vector and raster datasets used in the map-type
figure example are taken.

The content of this document is similar to an R Markdown document but in contrast
to R Markdown, shiny requires either an active R session or a shiny server to 
be served online.

Basically, all you have to do is to come up with standard R code. The part of the
code which should be adapted by some interaction has to be included inside the
`renderPlot` function. To change that code, include one or more input selectors
inside the `inputPanel` function. The actual value of the variables changed
by the input selectors can be accessed by calling the column of a data frame 
called "input" which has the same name as the keyword used for the input selector.

For more information, have a look at the [shiny documentation](http://rmarkdown.rstudio.com/authoring_shiny.html).

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
if(Sys.info()["user"] == "shiny"){
  in_path1 <- "/srv/shiny-server/moc/msc-data-management/data_procd/"
  in_path2 <- "/srv/shiny-server/moc/msc-data-management/data_raw/"
  in_path3 <- "/srv/shiny-server/moc/msc-data-management/data_procd/spatial_data/"
  } else {
    setwd("D:/active/moc/dm/examples/")
    in_path1 <- paste0(getwd(), "/data_procd/")
    in_path2 <- paste0(getwd(), "/data_raw/")
    in_path3 <- paste0(getwd(), "/data_procd/spatial_data/")
    }

packages <- c("raster", "rgdal", "latticeExtra", "caret", "gridExtra")
lapply(packages, library, character.only = TRUE)

wb <- read.table(paste0(in_path1, "wb-db_gnipc_co2_1960-2013.csv"), 
                 header = TRUE, sep = ",")
cntry <- read.table(paste0(in_path2, "wb-db_country_classes.csv"), 
                    header = TRUE, sep = ",")
wbc <- merge(wb, cntry[,!(grepl("Economy|X", colnames(cntry)))], 
             by.x = "Country.Code", by.y = "Code")
wbc$Region[wbc$Region == ".."] <- NA
wbc$Region <- droplevels(wbc$Region)
wbc$co2_log <- log(wbc$co2)
wbc$gni_log <- log(wbc$gni)
wbc <- wbc[complete.cases(wbc),]
```

##### Histogram
The first example shows two histogram plots which are plotted next to each other
by using `grid` function from the `gridExtra` library.

The `inputPanel` function contains two input selectors: a drop down menue
defined by the `selectInput` function and a slider defined by the `sliderInput`
function. Each of those input selectors is identified by a keyword (inputId) and
the actual value of the selector can be accessed by a data frame called "input".
This data frame is constructed by shiny in the background and can not be seen
in the code. For example, to access the content of the drop down menue, just
use "input$n_breaks".

The `renderPlot` function contains the two plots (one for GNI, one for CO2).
Please note that this is standard R code (i.e. lattice plot code) so there is
nothing new. First, the histogram of the GNI values is stored in a variable
called "hist1". Second, the density line is stored in a variable "dens1" and 
afterwards both are combined into one plot stored "plot1". The same is done
for CO2 afterwards and finally both plots are visualized by the 
`grid.arrange` function.

The only thing which is not standard inside the `renderPlot` function is the 
usage of the "input" data frame since it has not been defined in the code but
which is defined by shiny as a result from the `inputPanel` function. That's all.

```{r, echo=FALSE, warning=FALSE, }
inputPanel(
  
  selectInput(inputId = "n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput(inputId = "bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2,
              animate = animationOptions(interval = 1000, loop = TRUE, 
                                         playButton = NULL, pauseButton = NULL))
  )

renderPlot({
  hist1 <- histogram(wbc$gni_log, type = "density", 
            breaks = as.numeric(input$n_breaks), col = "white",
       xlab = "Logarithm of GNI", ylab = "Density", 
       main = "World bank data GNI values")

  dens1 <- densityplot(wbc$gni_log, adjust = input$bw_adjust, na.rm = TRUE, 
              col = "darkgreen", lwd = 3.0, plot.points = FALSE)
  
  plot1 <- hist1 + as.layer(dens1)

  hist2 <- histogram(wbc$co2_log, type = "density", 
            breaks = as.numeric(input$n_breaks), col = "white",
       xlab = "Logarithm of CO2", ylab = "Density", 
       main = "World bank data CO2 values")

  dens2 <- densityplot(wbc$co2_log, adjust = input$bw_adjust, na.rm = TRUE, 
              col = "blue", lwd = 3.0, plot.points = FALSE)
  
  plot2 <- hist2 + as.layer(dens2)
  
  grid.arrange(plot1, plot2, ncol = 2)
  })
```


##### XYPlot

The next example shows the correlation between the CO2 and GNI as a scatter
plot. This time, the interactive component allows the selection of
different models which fit the data set.

Compared to the example above, only one input selector (drop down menue) is
used. As you can see, the basic structure stays the same and only the behaviour
(i.e. content) of the input selector (function `inputPanel`) and the code inside the `renderPlot` function is changed. The latter is standard R code which has nothing to do with shiny but just encompasses a couple of else if statements to
run the model which has been chosen by the use in the drop down menu as well as
the respective lattice plot routine for scatter plots (i.e. xyplot).

```{r, echo=FALSE, warning=FALSE, }
inputPanel(
  
  selectInput(inputId = "model_select", label = "Select model:",
              choices = c("Linear model", "Exponential model",
                          "Self starting logistic model", "Tolerance curve"),
              selected = "Linear model")
  )

renderPlot({

  if(input$model_select == "Linear model"){
    # linear model
    l_model <- lm(wbc$co2_log ~ wbc$gni_log)
    thv <- predict.lm(l_model, wbc)
    } 
  else if(input$model_select == "Exponential model"){
    f <- function(y0, x, a, b) {y0 + a * exp(b * x)}
    exp_model <- nls(co2_log ~ f(y0, gni_log, a, b),
                     data = wbc,
                     start = list(y0 = 3, a = 0.5, b = 0.5)) 
    thv <- predict(exp_model, wbc$gni_log)
    } 
  else if(input$model_select == "Self starting logistic model"){
    Asym <- 1; xmid <- 1; scal <- 1
    getInitial(co2_log ~ SSlogis(gni_log, Asym, xmid, scal), data = wbc)
    ssl_model <- nls(co2_log ~ SSlogis(gni_log, Asym, xmid, scal), data = wbc)
    thv <- predict(ssl_model, wbc$gni_log)
    } 
  else if(input$model_select == "Tolerance curve"){
    tc_model= nls(co2_log ~ 1 + a * (gni_log^2) + b * gni_log + c, wbc,
                  start = list(a=1, b=1, c=1))
    thv <- predict(tc_model, wbc$gni_log)
    }
  
  xyplot(wbc$co2_log[wbc$co2_log >  thv] ~ 
           wbc$gni_log[wbc$co2_log >  thv],
         col = "red", ylim = range(wbc$co2_log, na.rm = TRUE),
         xlab = "Logarithm of GNI", ylab = "Logarithm of CO2", 
         main = "World bank data values") +
    as.layer(xyplot(wbc$co2_log[wbc$co2_log <=  thv] ~ 
                      wbc$gni_log[wbc$co2_log <=  thv], 
                    col = "blue")) + 
    as.layer(xyplot(thv ~ wbc$gni_log, type = "spline", 
                    col = "green", lwd = 5))
  })
```



##### Random forest
This final example uses a pre-computed random forest model which is loaded into
R using the `load` function. The only difference to standard R code is again the
drop down menue defined in the `inputPanel` function and the wrapping of those
parts of the code which actually consider the user input inside the `renderPlot`
function. 

```{r, echo=FALSE, warning=FALSE, }
load(paste0(in_path3, "dm-es-12_02_rf.RData"))
ndvi_original <- raster$ndvi_fogo_landsat
prediction_original <- predict(raster, model.train[[1]])
clrs.ndvi <- colorRampPalette(brewer.pal(8,"YlGn"))(256)
clrs.change <- colorRampPalette(brewer.pal(11,"Spectral"))(256)

inputPanel(
  selectInput(inputId = "ndvi", label = "Change percentage NDVI:",
              choices = seq(0, 200, 10), selected = 100)
  )

renderPlot({
  raster$ndvi_fogo_landsat <- ndvi_original * (as.numeric(input$ndvi) / 100)
  prediction_raster <- predict(raster, model.train[[1]])

  
  plot1 <- spplot(prediction_raster,col.regions = clrs.ndvi,
                  colorkey = list(at = seq(0, 11, length.out = 256)),
         main = "Predicted species richness \n based on Landsat 8 data",
         sub = paste0("Result is based on ",input$ndvi, "% of 2014 NDVI values"),
         scales = list(draw = TRUE))
  change_raster = (1-prediction_raster / prediction_original) * 100
  plot2 <- spplot(change_raster,col.regions = clrs.change, 
                  colorkey = list(at = seq(-100, 100, length.out = 256)),
                  main = "Percentage change in prediction \n results compared to 100% NDVI",
                  scales = list(draw = TRUE))
  grid.arrange(plot1, plot2, ncol = 2)
  })
```
