<!---
Version: "2013-10-23"
Copyright (C) 2013 Thomas Nauss, GNU GPL (>=3)
-->
#### Readading data from csv files

Reading csv files is realized using the `read.table` function from R's utils
library. The function will return a data frame which contains the information of
the csv file.

```{r}
df <- read.table("D:/active/moc/dm/examples/data_raw/wb-db_co2em_1960-2013.csv",
                 header = TRUE, sep = ",", skip = 2)
```

As you can see, the `read.table` function gets several arguments. Since this
is probably the first time you use an R function, let's have a closer look on
these arguments. The first one gives the filename inclducing the path to the 
file starting from the level set as working directory. The second argument
`header = TRUE` tells the function, that the csv file has a header line which is 
used by `read.table` to name the columns of the returning data frame. The third 
one `sep = ","` defines the separator of the individual columns in the data 
frame. Finally, since our csv file has an additional header and an empty line 
below (in lines 1 and 2), the function should skip the first two lines 
(`skip = 2`).

A note on the sequence of the arguments: the sequence of the arguments does
not matter as long as you name them explicetly. If you do not use the argument
identfier as it is the case for the first argument, the filename, in the example 
then the sequence matters. To get information on the default sequence and of 
course the general application of the each R function, type `?<function name>`
(e.g. `?read.table`) in an R console.

After you executing the `read.table` function above, the content of the csv file is
stored into a two dimensional data frame called `df`. 

A quick way to check if everything is fine is to display the first few lines of
the data file using the `head` function.
```{r}
head(df,2)
```

#### Writing data to csv files
Writing data is as easy as reading it. Just use the `write.table` function.
```{r}
write.table(df, file = "D:/active/moc/dm/examples/data_procd/test.csv", 
            sep = ";", dec = ",")
```
The example above also illustrates another argument called `dec` which defines
the decimal point. In addition, the parameter for the `sep` argument is ";"
which causes the new csv file to have ";" instead of "," as a separator. This
would be a good idea for German csv files which use the "," as decimal sign but 
to put this straight: CSV means comma separated values so just use "," as a 
separator and "."  as decimal point. 

For more information have a look at e.g. the respective [data importing](http://www.statmethods.net/input/importingdata.html) and [data exporting](http://www.statmethods.net/input/exportingdata.html) site at Quick R, have a look into the package 
documentation or search the web.